Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-06-15 ms.tex
various relatively minor edits

In $\S$\ref{sec:data} we introduce the data used throughout this study. In 
$\S$\ref{sec:prepro} we describe our preprocessing of the \PLANCK~maps to 
isolate thermal emission from Galactic dust. In $\S$\ref{sec:modeling} we 
explain the dust emission model we fit to the \PLANCK~and \IRAS~data. In 
$\S$\ref{sec:global} we describe our constraints on this model's global 
parameters derived from the \PLANCK~High Frequency Instrument (HFI) maps. In 
$\S$\ref{sec:fitting} we detail the Markov chain Monte Carlo (MCMC) method with
which we have estimated the spatially varying parameters of our model. In 
$\S$\ref{sec:ebv} we calibrate our derived optical depth to optical reddening. 
In $\S$\ref{sec:results} we present the full-sky maps of dust temperature and 
optical depth we have obtained, and conclude in $\S$\ref{sec:conclusion}.

@@

\subsection{CMB Anisotropy Removal}
\label{sec:cmb}
We first addressed the CMB anisotropies before performing any of the 
interpolation/smoothing described in $\S$\ref{sec:ptsrc}/$\S$\ref{sec:smth}. 
The CMB anisotropies are effectively imperceptible upon visual inspection 
in \PLANCK~857 GHz, but can be perceived at a low level in \PLANCK~545 GHz, and
are prominent at 100-353 GHz relative to the Galactic emission
we wish to characterize, especially at high latitudes. To remove the CMB 
anisotropies, we have subtracted the Spectral Matching Independent Component 
Analysis \citep[SMICA,][]{smica} model from each of the \PLANCK~maps, 
after applying appropriate unit conversions for the 545 and 857 GHz maps 
with native units of MJy/sr. Low-order corrections, particularly removal of the
residual Solar dipole, are discussed in $\S$\ref{sec:zp}.

@@

Specifically, we perform a linear fit to the FDS99-predicted 857 GHz values as 
a function of \PLANCK~857 GHz. For this purpose, we employ a version of the 
\PLANCK~857 GHz map with zodiacal light and point sources removed and smoothed 
to 1$^{\circ}$ FWHM, which we will refer to as $\mathcal{I}_{857}$. We 
consider $\mathcal{I}_{857}$ to be the independent variable, as it has much 
higher S/N than the FDS99 prediction, henceforward referred to as 
$\mathcal{F}_{857}$. Note that $\mathcal{F}_{857}$ is not simply the FDS99
model evaluated at 857 GHz, but also incorporates the color correction factor 
of $\S$\ref{sec:bpcorr}, using the FDS99 temperature map to determine the dust 
spectrum shape. We then rebin to $N_{side}$=64 and restrict to pixels with 
$\mathcal{I}_{857}$$<$2.15 MJy/sr. Since \PLANCK~857 GHz smoothed to degree 
resolution has very high S/N, we can safely perform such a cut on 
$\mathcal{I}_{857}$. Figure \ref{fig:fdsref} shows a scatterplot of 
$\mathcal{I}_{857}$ versus $\mathcal{F}_{857}$, with a moving median and linear
fit overplotted. The linear fit was performed with uniform weights and 
iterative outlier rejection. The best-fit linear model is given by 
$\mathcal{F}_{857}$=0.991$\mathcal{I}_{857}$$-$0.018. It is encouraging that 
the slope is quite close to unity, and the offset near zero.

@@

The formal statistical error on the best-fit 857 GHz offset is quite small, 
$\sim$0.002 MJy/sr. The systematics likely to dominate the actual uncertainty 
on our FDS-based zero level are imperfections in the \PLANCK/\verb|i100| 
zodiacal light models and the FDS99 temperature map. To quantify these 
systematic uncertainties, we split the sky into four quadrants, with the 
dividing boundaries being $b$=0$^{\circ}$ and $l$=0$^{\circ}$, 
$l$=180$^{\circ}$. We again restricted to $\mathcal{I}_{857}$$<$2.15 MJy/sr, 
and repeated the regression in each quadrant. The rms of the per-quadrant 
slopes was found to be 0.0188, while the rms of the per-quadrant offsets was 
0.0586 MJy/sr.

@@

In the course of this study we use not only \PLANCK~857 GHz, but also all of
the remaining \PLANCK~HFI bands, as well as \verb|i100|. To derive the 
zero level offsets that must be applied to each of the five lowest-frequency 
\PLANCK~bands, we perform a regression versus the \PLANCK~857 GHz map corrected
for the absolute zero level offset derived in $\S$\ref{sec:zp_abs}. We assume 
no offset need be applied to \verb|i100|, which already has its zero level 
tied to H\,\textsc{i}.

The need for additional low-order corrections beyond simple scalar offsets 
became evident upon inspecting the HFI maps at 100-545 GHz. In particular, we 
noticed the presence of a low-level dipole pattern, with an orientation 
consistent with that of the Solar dipole. Our strategy will be to 
simultaneously fit both the residual dipole and zero-level offset amplitudes. 
To most precisely recover these amplitudes, it is necessary to have the highest
available S/N in the independent variable of our regression. For this reason we
have used 857 GHz as a reference for the 100-545 GHz bands, as opposed to the 
FDS99 predictions or H\,\textsc{i} data.

Thus, we perform one regression per HFI band to simultaneously fit for the 
zero level offset, the slope relative to 857 GHz, and the residual dipole 
amplitude. For each 100-545 GHz HFI band, we restrict to regions of low column 
density (H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for $-72$$<$$v_{LSR}$$<$$+25$ 
km\,s$^{-1}$) and fit the following model:

@@

With $p$ denoting a single nside=64 HEALPix pixel \citep{healpix} in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{\nu_i}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and the constant offset of $\S$\ref{sec:zp_abs} removed, smoothed to 
$1^{\circ}$ resolution. $\mathcal{I}_{\nu_i}$ is the corresponding $1^{\circ}$ 
resolution \PLANCK~HFI map with zodiacal emission and compact sources removed. 
In the context of Equation \ref{equ:dip}, $\nu_i \in$  
\{100, 143, 217, 353, 545\} GHz. Note that $\mathcal{I}_{\nu_i}$ is always in
the native units of the relevant \PLANCK~band. $\mathcal{D}$ is a scaling of 
the Solar dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$, with unit amplitude. Because 
$\sim$18,000 pixels satisfy the low H\,\textsc{i} cut, we have an 
overconstrained linear model with three parameters: $m$, $d$, and $b$. $m$ 
represents the best-fit slope of \PLANCK~band $\nu_i$ versus \PLANCK~857 GHz 
assuming they are linearly related. $d$ is the residual Solar dipole amplitude,
and its best-fit value represents the scaling of the Solar dipole that makes 
the \PLANCK~band $\nu_i$ versus 857 GHz correlation most tightly linear. $b$ 
represents the constant offset that must be subtracted from the band $\nu_i$ 
map to make its zero level consistent with that of the 857 GHz map. For each 
band $\nu_i$, we obtain estimates of $m$, $d$, and $b$ by performing a linear 
least squares fit with uniform weights and iterative outlier rejection. Figure 
\ref{fig:dip} shows scatterplots of the band $\nu_i$ versus 857 GHz correlation
 before and after correcting for the best-fit offset and residual dipole, for 
each $\nu_i \in$ \{143, 217, 353, 545\} GHz. Not only are the tightened 
correlations striking in these scatterplots, but the residual dipole 
subtractions appear very successful in the two-dimensional band $\nu_i$ maps 
themselves. Before performing thermal dust fits, we therefore subtract the 
best-fit $b$ and $d\mathcal{D}$ from each band $\nu_i$ map. The best-fit 
offsets and residual dipole amplitudes are listed in Table \ref{table:offs},
along with other important per-band parameters, such as the fractional 
multiplicative calibration uncertainty $c_{\nu}$.

>>>>

In $\S$\ref{sec:data} we introduce the data used throughout this study. In 
$\S$\ref{sec:prepro} we describe our preprocessing of the \PLANCK~maps to 
isolate thermal emission from Galactic dust. In $\S$\ref{sec:modeling} we 
explain the dust emission model we fit to the \PLANCK~and \IRAS~data. In 
$\S$\ref{sec:global} we describe our constraints on this model's global 
parameters derived from the \PLANCK~High Frequency Instrument (HFI) maps. In 
$\S$\ref{sec:fitting} we detail the Markov chain Monte Carlo (MCMC) method with
which we have estimated the spatially varying parameters of our model. In 
$\S$\ref{sec:ebv} we calibrate our derived optical depth to reddening at 
optical wavelengths. In $\S$\ref{sec:results} we present the full-sky maps of 
dust temperature and optical depth we have obtained, and conclude in 
$\S$\ref{sec:conclusion}.

@@

\subsection{CMB Anisotropy Removal}
\label{sec:cmb}
We first addressed the CMB anisotropies before performing any of the 
interpolation/smoothing described in $\S$\ref{sec:ptsrc}/$\S$\ref{sec:smth}. 
The CMB anisotropies are effectively imperceptible upon visual inspection 
of \PLANCK~857 GHz, but can be perceived at a low level in \PLANCK~545 GHz, and
are prominent at 100-353 GHz relative to the Galactic emission
we wish to characterize, especially at high latitudes. To remove the CMB 
anisotropies, we have subtracted the Spectral Matching Independent Component 
Analysis \citep[SMICA,][]{smica} model from each of the \PLANCK~maps, 
applying appropriate unit conversions for the 545 and 857 GHz maps with native 
units of MJy/sr. Low-order corrections, particularly removal of the residual 
Solar dipole, are discussed in $\S$\ref{sec:zp}.

@@

We perform a linear fit to the FDS99-predicted 857 GHz values as 
a function of \PLANCK~857 GHz. For this purpose, we employ a version of the 
\PLANCK~857 GHz map with zodiacal light and point sources removed and smoothed 
to 1$^{\circ}$ FWHM, which we will refer to as $\mathcal{I}_{857}$. We 
consider $\mathcal{I}_{857}$ to be the independent variable, as it has much 
higher S/N than the FDS99 prediction, henceforward referred to as 
$\mathcal{F}_{857}$. Note that $\mathcal{F}_{857}$ is not simply the FDS99
model evaluated at 857 GHz, but also incorporates the color correction factor 
of $\S$\ref{sec:bpcorr}, using the FDS99 temperature map to determine the dust 
spectrum shape. We then rebin to $N_{side}$=64 and restrict to pixels with 
$\mathcal{I}_{857}$$<$2.15 MJy/sr. Since \PLANCK~857 GHz smoothed to degree 
resolution has very high S/N, we can safely perform such a cut on 
$\mathcal{I}_{857}$. Figure \ref{fig:fdsref} shows a scatterplot of 
$\mathcal{I}_{857}$ versus $\mathcal{F}_{857}$, with a moving median and linear
fit overplotted. The linear fit was performed with uniform weights and 
iterative outlier rejection. The best-fit linear model is given by 
$\mathcal{F}_{857}$=0.991$\mathcal{I}_{857}$$-$0.018. It is encouraging that 
the slope is quite close to unity, and the offset near zero.

@@

The formal statistical error on the best-fit 857 GHz offset is quite small, 
$\sim$0.002 MJy/sr. The systematics likely to dominate the actual uncertainty 
on our FDS-based zero level are imperfections in the \PLANCK/\verb|i100| 
zodiacal light models and the FDS99 temperature map. To quantify these 
systematic uncertainties, we split the sky into four quadrants, with boundaries
at $b$=0$^{\circ}$ and $l$=0$^{\circ}$, $l$=180$^{\circ}$. We again restricted 
to $\mathcal{I}_{857}$$<$2.15 MJy/sr, and repeated the regression in each 
quadrant. The rms of the per-quadrant slopes was found to be 0.0188, while the 
rms of the per-quadrant offsets was 0.0586 MJy/sr.

@@

In the course of this study we use not only \PLANCK~857 GHz, but also all of
the remaining \PLANCK~HFI bands, as well as \verb|i100|. To derive the 
zero level offsets that must be applied to each of the five lowest-frequency 
\PLANCK~bands, we perform a regression versus the \PLANCK~857 GHz map corrected
for the best-fit absolute zero level offset from $\S$\ref{sec:zp_abs}. We 
assume no offset need be applied to \verb|i100|, which already has its zero 
level tied to H\,\textsc{i}.

The need for additional low-order corrections beyond simple scalar offsets 
became evident upon inspecting the HFI maps at 100-545 GHz. In particular, we 
noticed the presence of a low-level dipole pattern, with an orientation 
consistent with that of the Solar dipole. Our strategy will be to 
simultaneously fit both the residual dipole and zero-level offset amplitudes. 
To most precisely recover these amplitudes, it is necessary to have the highest
available S/N in the independent variable of our regression. For this reason we
have used \PLANCK~857 GHz as a reference for the 100-545 GHz bands, as opposed 
to the FDS99 predictions or H\,\textsc{i} data. In doing so, we assume
\PLANCK~857 GHz contains no appreciable Solar dipole residual or other 
low-order problems. % grammar?

We perform one regression per HFI band (other than 857 GHz) to simultaneously 
fit for the zero level offset, the slope relative to 857 GHz, and the residual 
dipole amplitude. For each 100-545 GHz HFI band, we restrict to regions of low 
column density (H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for 
$-72$$<$$v_{LSR}$$<$$+25$ km\,s$^{-1}$) and fit the following model:

@@

With $p$ denoting a single nside=64 HEALPix pixel \citep{healpix} in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{\nu_i}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and the constant offset of $\S$\ref{sec:zp_abs} removed, smoothed to 
$1^{\circ}$ resolution. $\mathcal{I}_{\nu_i}$ is the corresponding $1^{\circ}$ 
resolution \PLANCK~HFI map with zodiacal emission, CMB anisotropies, and
compact sources removed. In the context of Equation \ref{equ:dip}, $\nu_i \in$ 
\{100, 143, 217, 353, 545\} GHz. Note that $\mathcal{I}_{\nu_i}$ is always in
the native units of the relevant \PLANCK~band. $\mathcal{D}$ is a scaling of 
the Solar dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$, with unit amplitude. Because 
$\sim$18,000 pixels satisfy the low H\,\textsc{i} cut, we have an 
overconstrained linear model with three parameters: $m$, $d$, and $b$. $m$ 
represents the best-fit slope of \PLANCK~band $\nu_i$ versus \PLANCK~857 GHz 
assuming they are linearly related. $d$ is the residual Solar dipole amplitude,
and its best-fit value represents the scaling of the Solar dipole that makes 
the \PLANCK~band $\nu_i$ versus 857 GHz correlation most tightly linear. $b$ 
represents the constant offset that must be subtracted from the band $\nu_i$ 
map to make its zero level consistent with that of the 857 GHz map. For each 
band $\nu_i$, we obtain estimates of $m$, $d$, and $b$ by performing a linear 
least squares fit with uniform weights and iterative outlier rejection. Figure 
\ref{fig:dip} shows scatterplots of the band $\nu_i$ versus 857 GHz correlation
 before and after correcting for the best-fit offset and residual dipole, for 
each $\nu_i \in$ \{143, 217, 353, 545\} GHz. Not only are the tightened 
correlations striking in these scatterplots, but the residual dipole 
subtractions appear very successful in the two-dimensional band $\nu_i$ maps 
themselves. Before performing thermal dust fits, we therefore subtract the 
best-fit $b$ and $d\mathcal{D}$ from each band $\nu_i$ map. The best-fit 
offsets and residual dipole amplitudes are listed in Table \ref{table:offs},
along with other important per-band parameters, such as the fractional 
multiplicative calibration uncertainty $c_{\nu}$.
